---
title: "Vaulting into Victory: Optimizing for US Gymnastics Medal Count at the Paris 2024 Olympics"
author: "Sean Li, Christopher Tsai, Benjamin Thorpe"
date: "2023-11-21"
geometry: "left=2cm,right=2cm,top=1.5cm,bottom=1.5cm"
output: 
  pdf_document:
    fig_caption: yes
warnings: false
fontsize: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

### Background

The Olympics brings about a sense of national pride seldom felt by athletes and viewers alike. The sheer amount of viewers adds to the pressure and importance of an event like this for national athletic committees. In every Olympic games since 2008, over 3 billion unique viewers tuned in worldwide to view the games far surpassing the 115 million viewers that tuned into the Super Bowl. Americans, in particular, are fascinated by artistic gymnastics with the highest percentage of those who watch the Olympics claiming to be interested in the event (Tracy, 2021). It follows that success in this sport has long been a source of international recognition for the United States. The breathtaking performances and athleticism displayed by American gymnasts have not only brought numerous Olympic medals home, but also served as inspiration for generations of aspiring athletes. Simone Biles, Gabby Douglas, and Nastia Liukin are all household names due to their success in Olympic gymnastics. As we approach the next Olympic Games in Paris 2024, it is vital that team USA finds success in the form of medals in artistic gymnastics. In the era of data-driven decision-making, the role of predictive analytics in sports has gained unprecedented significance and the need for innovation and precision in athlete selection has never been more crucial. Our goal is to use data analytics to predict the best five-member lineups for both the USA Men’s and Women’s artistic gymnastics teams, optimizing for total medal count while taking into account the weight placed on team success.

First, understanding Olympic gymnastics and its scoring system will be vital to understanding the decision making process later in this paper. Women’s artistic gymnastics and Men’s artistic gymnastics vary slightly, with women competing in four apparatuses and men competing in six. Each country, if they have qualified for the team all-around competition (which the US has), can send five athletes per gender; otherwise a country can send a maximum of three individuals. The event begins with a qualification round that is also the qualifying round for the team all-around, individual all-around, and individual apparatus competitions. A team will send four of their five athletes to compete for each of the apparatuses. For the team score, only the top three of the four scores will count. If an individual competes on every apparatus in the qualifying round they are then eligible for the individual all-around final, regardless of if they are on a team. However, only the top 24 athletes will be in the individual all-around final. If an individual is in the top eight in qualifying for a given apparatus they also qualify for that individual apparatus final. For all individual finals, however, only a maximum of two athletes per country can qualify. All things considered, this gives six medal opportunities on the women’s side (team all-around, individual all-around, each of the four apparatuses) and eight on the men’s side (team all-around, individual all-around, each of six apparatuses).

This report aims to analyze athlete-specific data and incorporate domain expertise from the gymnastics community. It will offer a comprehensive and customized approach that addresses the unique competition format presented by artistic gymnastics and provides the USA Olympics Committee with a robust tool to make statistically informed decisions. Our goal is to build a tool that will allow us to recommend a team of five women and five men for the Olympic Games in Paris 2024, with an expected individual medal count for each team as well as an expected all-around team medal count given the five person team.

### Exploratory Data Analysis

If our goal is to put forth the best lineup to maximize medal count, we need to look into who the top US candidates are in each apparatus and all-around. Here, we graph dot plots of the top three US gymnasts in each event across all the events, using the mean score from their 2022-2023 data. The small number of total athletes that appear in the top three across all apparatuses signals that most elite athletes are great at more than one event.

```{r, echo=FALSE, results = FALSE, warning= FALSE, message=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(stringr)
library(lubridate)
library(knitr)
```

```{r data cleaning, echo=FALSE, results = FALSE, warning= FALSE, message=FALSE}
#| label: load-data
data_2022_2023 <- read_csv("App-1/data/cleaned.csv")
```

```{r usa-apparatus-means, echo=FALSE}
data_2022_2023 <- data_2022_2023 %>% 
  mutate(Name = paste(FirstName, LastName))

top5usa <- function(gen, ap){
  df <- data_2022_2023 %>%
    filter(Gender == gen, Country == "USA", Apparatus == ap) %>% 
    group_by(Name) %>%
    summarise(mean = mean(Score), samples = n()) %>%
    filter(samples >=6) %>% 
    arrange(desc(mean)) %>%
    top_n(3, mean) %>%
    inner_join(data_2022_2023 %>% filter(Apparatus == ap), by = c("Name"="Name")) %>%
    arrange(desc(mean)) %>% 
    select(Name, Apparatus, Score, DaysUntil2024, mean) %>%
  return(df)
}
usawomenbb <- top5usa("w", "BB")
topusawomen <- rbind(top5usa("w", "BB"),top5usa("w","UB"), top5usa("w","FX"), top5usa("w","VT")) 
topusmen <- rbind(top5usa("m", "HB"),top5usa("m","PB"), top5usa("m","FX"), top5usa("m","VT"), top5usa("m","PH"), top5usa("m","SR")) 
```

```{r plotuswomen, echo=FALSE}
u5w <- ggplot(topusawomen, aes(y = reorder(Name,mean), x = Score, color = Name)) + 
  geom_point(alpha = 0.5) +
  facet_wrap(~Apparatus, drop = TRUE) + 
  labs(title = "Top 3 USA gymnastics women per apparatus", subtitle = "Biles appears in top 3 across all 4 apparatus",
       y = " " )+ theme_bw()
u5w
```

Taking a glance at the top US women, we can see that Simone Biles, who is a top three gymnast in each of the four apparatuses, is by far the best American gymnast. That is unsurprising given that she has the most gymnastics medals in the world and has consistently performed at an elite level since her return to the sport. She is a lock to make the Paris 2024 roster. Shilese Jones is the only gymnast to have a higher mean score in an apparatus than Simone, outperforming her in the uneven bars (UB). She is another athlete to keep an eye out for to make the team, especially considering her recent all-around Bronze medal in the World Championships in October 2023.

```{r usmen, echo=FALSE}
u5m <- ggplot(topusmen, aes(y = reorder(Name,mean), x = Score, color = Name)) + 
  geom_point(alpha = 0.5) +
  facet_wrap(~Apparatus, drop = TRUE) + labs(title = "Top 3 USA gymnastics men per apparatus", subtitle = "Juda, Young, Hong appear in top 3 in half of the apparatus'",
       y = " ") + theme_bw() + theme(axis.text.y.left = element_text(size=7.5), legend.position = "none")
u5m
```

Paul Juda, Khoi Young, and Asher Hong appear in the top 3 gymnasts for half of the apparatus although determining the US men’s team is not quite as clear cut. There are a total of 11 American gymnasts who placed in the top three of at least one of the six events. Hong, Young, and Yul Moldauer placed in the top three for three of the six events. Additionally, Brody Malone placed in the top five for four out of six events, and Fred Richard most recently won Bronze at the 2023 World Championships in the all round. Those are athletes we might expect to be placed into our US men’s team. 

## Methodology

### Weighting

We treated this competition as an optimization problem where we maximize total medal count. We have developed nuanced heuristics based on analyzing past Olympic gymnastic results to pick our top five-person lineups. Specifically, we weighted for each score by artificially determining its importance by how recent it was and the “stage” of the competition in which it took place, qualifying vs. final round. We averaged the weighted scores across all the events a US gymnast has participated in since the 2021 Tokyo Olympics. A composite all-around score was found as well by summing the average weighted scores in each apparatus. A summary of our weighting process is below.

### Clutch Factor and Recency Considerations
We chose to put weights on competition date and competition stage, specifically, because we believe a heavier weight should be placed on performances that take place under pressure (the final round) and are more recent. This weighting attempts to measure the “clutch” factor and peak/prime performance of an individual athlete. We duplicate rows of individual scores in the sampling dataset given the assigned weight of that score. Competition date will be split into three categories by recency: very recent (< 6 months), somewhat recent (6 - 18 months), not recent (18+ months). A multiplier of one will be given to ‘not recent’ competitions, a multiplier of two to ‘somewhat recent’ competitions, and a multiplier of three for ‘very recent’ competitions. As for the competition stage, there are two stages of competition: qualifiers and finals. Qualifying round scores will receive a multiplier of one and final round scores will receive a multiplier of two. For example, a row containing a score from a not recent competition and qualifying round will remain as one singular row. However, a row containing a qualifying round score, but from a very recent competition, will be multiplied three times to get three rows containing that score. If that score was instead from the finals, it would be multiplied 2 x 3 = 6 times in the dataset. Taking the average of an individual athlete’s scores after this duplication process will give us a weighted mean expected score on each apparatus for that athlete. Still, each gymnast has variance across performances on each apparatus which is taken into account. For each USA gymnast, we developed and will sample from an individual total score distribution from their weighted scores to account for the variance for each individual on each apparatus. A density plot of all the scores after the duplication process will give us a distribution centered around the mean score for an athlete as well as their variance.

### Obtaining Medaling Probabilities
The main goal of analysis is to determine the probability that an individual athlete medals, so we also had to determine a way to get medal probabilities. We accomplished this goal by determining score thresholds necessary to achieve a certain medal. This first required us to get a distribution of scores on individual apparatuses. From the EDA, we saw that the distribution of all scores was roughly normal, with a tail towards the left side of the distribution, with fatter tails. We also saw that each individual athlete’s score distribution was roughly normal. This inspired us to take the top 24 people (the lower scoring competitors would be dropped), and from there create a t-distribution for the 24 people as well as for each individual competitor to sample from in our simulation. This will account for the individual variance of athletes on a specific apparatus. Of the 24 gymnasts, however, no more than four competitors from the same country will be included in this distribution as a maximum of four individuals from each country can compete in qualifying for each apparatus. In order to find an individual’s probability of earning a medal we used a two step simulation and sampling process. We simulated the following process 1,000 times. We first sampled from the distribution of the top 24 individuals 24 times. The top three scores of the 24 represented the three medal cutoffs with the 3rd highest sample score corresponding to a Bronze medal, the 2nd highest sample score corresponding to a silver medal cutoff, and the highest sample score corresponding to a gold medal cutoff. Because we were only worried about gymnasts from the USA, we sampled once from each of the top eight USA gymnasts in each individual apparatus given their unique distributions. If that gymnast’s sampled score surpassed one of the cutoffs they ‘earned’ that medal for the given simulation. For example, if when sampling from the 24 gymnasts distribution we get medal cutoffs of 14.5, 14.8, and 14.9 and gymnast A gets a sampled score of 14.5 or greater they “earn” a Bronze medal. A sampled score of 14.8 or greater would “earn” a Silver medal and a sampled score of 14.9 or greater would “earn” a Gold medal. The only caveat was if multiple American gymnasts fall within the threshold for a certain medal. In that case the gymnasts were ordered and the top score earned that medal and the next highest score would earn the next highest medal if available. We simulated this approach 1,000 times to get each individual gymnast’s probability of getting a specific medal in a specific apparatus. We used this approach for the individual all-around as well. The assumptions for our simulation require an adequate amount of repeating sampling. 1,000 is surely enough for each athlete, who on average has 10-20 observations in our dataset. We also assume that past performance is somewhat indicative of future performance. 

From there we used optimization to determine the 5-person lineup that maximizes expected medals earned including the team all-around medal. In determining the team all-around medal we will use a system identical to the way the Olympics will work in 2024. We will use many permutations of five gymnasts and sample from every gymnast’s individual distributions to get all-around scores. For each apparatus, the top three US gymnasts on the team’s distributions will be sampled from (because only the top three gymnasts participate in the team all-around finals). The scores will be aggregated to get a total team all-around score. The same method of medal cutoffs will be used in the team all-around medal cutoffs, but using empirical team all-around score data to create a distribution to be sampled from.

### Our Web App
Our web app then combines the methodology above into a single interface for athlete selection and optimization. There are three tabs of note within this web app. First, an about section which describes the creators of this app as well as a short synopsis of how the web app was created. Second, a “Gymnast” tab which allows the user to sort in descending order not only the projected total medals by athlete, but also for medal probability on individual apparatuses. This tab allows the user to see which US gymnasts (men or women) are the most likely to medal in a given apparatus. Finally, there’s the “Team Builder” page where for both the men and women’s teams, users will be able to manually input five gymnasts into the web app. Utilizing the several already run simulations, the app will spit out the total projected medals for each athlete, their all-around team medal probability, and finally the total projected combined medals for the five athletes. Replacing one athlete with another will also replace that athletes projected individual medals as well as recalculate that five-person team’s team all-around medal probability.

## Results / Our Team

```{r results, echo=FALSE}
results <- data.frame(
  Category = c("All-Around", "All-Around", "Specialist", "Specialist", "Specialist", "Total Exp. Medals"),
  USA_Women = c( "Simone Biles", "Shilese Jones", "Kaliya Lincoln", "Jade Carey", "Zoe Miller", ""),
  EMC_Women = c('3.409', '0.466', '0.439','0.190','0.061','4.565'),
  USA_Men = c("Paul Juda", "Fred Richard", "Donnell Whittenburg", "Khoi Young", "Curran Phillips", ""),
  EMC_Men = c('0.663', '0.491', '0.264', '0.564', '0.310',"2.402")
)
kable(results, caption = "Proposed Paris 2024 USA Gymastics Team, with Expected Medal Counts (EMC)")
```

Based on our results and our weights, to maximize for total medal count at the Paris 2024 Olympics, the USA Gymnastics team should select Simone Biles, Skye Blakely, Shilese Jones, Jade Carey, and Zoe Miller for the women’s team, and Fred Richard, Paul Juda, Donnell Whittenburg, Khoi Young, and Curran Phillips for the men’s team. Our total predicted medal counts are 4.565 individual medals for the women, and 2.402 for the men’s team. Biles and Jones are the clear all-round selections, expecting to contribute 3.409 and 0.466 medals in total, respectively, and having a 99.9% and 23.9% to medal in the individual all-around finals. Onto the three specialists, Carey is an outstanding performer in vault and floor exercise, having a 19.0% and 19.9%, respectively, to medal in those apparatuses. Lincoln has a 43.9% chance medaling in floor exercise. Miller is our last selection, and she brings value in the uneven bars, having a 6.1% percent chance of medaling there. Juda and Richard were the all-round selections on the men’s side. They have the highest chances of medaling in our individual all-round, with 5.1% and 4.9%, respectively. The first specialist we choose is Khoi Young for his ability in vault and pommel horse, having a 44.9% and 11.5%, respectively, of medaling in those events. Then we choose the veteran Whittenburg for his consistency on the still rings (SR), having the highest percentage chance to help the US medal in that apparatus where we expect him to come in the top three 26.4% of the time. Lastly, we went with Curran Phillips for parallel bars as he has a 26.3% to medal there.

While the team above maximizes the total projected amount of medals earned at the 2024 Paris Olympics. If instead, a user wanted to maximize the likelihood of winning a team medal, they could primarily focus on building permutations of teams in our “team builder” feature that maximize the probability of winning Gold, Silver, Bronze or any of the three based on user preference.

## Discussion

### Goals, Eye-test, and Outside Research
Our goal was to design a tool to recommend a team of five women and five men for the Olympic Games in Paris 2024, and to come up with an expected medal count for each team. We accomplished our goal of selecting athletes by first adjusting for ‘clutch’ factor and adding a recency bias by duplicating observations based on final vs qualifying rounds, and date of competition. We then account for individual variance by creating a t-distribution for each competitor for each apparatus and sampling from that distribution 1,000 times. We also find the predicted medal thresholds using repeated sampling from a t-distribution, taking the top 24 individual scorers, and comparing our sampled individual results with these predicted thresholds to find predicted individual medal counts. 

### Excluded Athletes

Athletes Konnor McClain, Sunisa Lee, and Brody Malone were excluded from our selections. Konnor McClain left Elite (which is a program designed to be a pathway to the USA national team) in 2023 in favor of college. Although she plans to continue training for the Olympics, her non-Elite status likely keeps her from being in the USA olympic committee’s player pool. A similar situation exists for Sunisa Lee, the 2020 Olympic all-around champion. Although her weighted scores are quite high, she has not competed in Elite since her enrollment in college in 2022 and a recent injury in which she gained over 50 pounds likely keeps her from making the USA national team. Injury concerns will also keep Brody Malone off the team on the men’s side. Although he eyes a spring 2024 comeback from a knee injury suffered in the spring of 2023, he will likely not be ready or in time for the Olympics.

### Limitations

There were some limitations identified in our statistical methods. First, we arbitrarily assigned weights to each score and it is possible that different weightings could result in the selection of a different team. However, the weights used were reasonable given how recent performances are more likely to affect future performance and the desire to pick ‘clutch’ gymnasts. Another potential limitation is how we determined projected scores for athletes that have not competed in some time. Take Sunisa Lee as an example. Due to competing in college and injuries, Lee has very few scores since her Olympic performance in 2021. Because those performances account for many of her competitions, her lower recent scores are overshadowed by her higher older scores. Even with our weighting scheme her older scores account for a much greater proportion of her data. In this vein, athletes may have projected scores that do not resemble what their recent form might be. Some alternatives for our methodology could include using a Kernel Density Estimation method to estimate the distribution and then sampling, or obtaining all possible combinations of five athletes and then running an Olympics simulation. We could also introduce more factors to weighting such as the age of the competitor, skill level of the competition, etc. 

### A Qualitative Analysis of Team USA

A qualitative analysis of Team USA’s women’s gymnastics team has determined that Team USA will most likely pick from a pool of seven women to choose their team, as only seven female gymnasts placed in the top three in any event in the most recent USA national gymnastics championships. These seven are Simone Biles, Shilese Jones, Leanne Wong, Skye Blakely, Kaliya Lincoln, Joscelyn Roberson, and Jade Carey. The two all-around staples are likely to be Simone Biles and Shilese Jones, as both placed first and second, respectively, in the USA championships as well as were the top two USA finishers in the World Championships. Leanne Wong has a case for being an all-around competitor as well, but in both championships she finished behind Biles and Jones. 

As for the specialists there is support for Leanne Wong being a third all-around athlete as this was how team USA utilized her in the World Championships. However, she did not qualify for any individual finals so her choice as such an athlete in the Olympics may not be the most optimal strategy. If we choose three specialists, Skye Blakely is a lock for the team. Not only did she place second in the vault, uneven bars, and balance beam, but she is by far and away the USA’s third best uneven bars athlete to go along with Biles and Jones. Beyond these three athletes, the biggest hole lies in the vault and floor exercise events. Of the three athletes remaining only Kaliya Lincoln fills the gap in the floor exercise as she not only won a silver medal at the national championships but also won gold at the Pan-American games. Kayla DiCello is another option in vault for the USA. While not as strong as Lincoln in this event, if the USA is looking for another more all-around option for added depth, Dicello provides just that, as she finished second in floor at the Pan-American games. Still, Kaliya Lincoln’s floor abilities would make her a good fourth addition. That leaves Joscelyn Roberson and Jade Carey, who are the two best options for the vault. Given her recent form, Roberson has a slight edge over Carey. Roberson won gold at the National Championships and had a good chance to medal at the World Championships before tweaking her left leg before the finals (which also prevented her from competing in the team all-around final). 

All in all, a qualitative analysis would suggest a five-woman team of Simone Biles, Shilese Jones, Skye Blakely, Kaliya Lincoln, and Joscelyn Roberson. Alternative lineups would feature either Kayla Dicello over Kaliya Lincoln or Jade Carey over Joscelyn Roberson.

A qualitative analysis for the Men’s Olympic team is much less clear. However, Frederick Richard is a lock for the team. He achieved a bronze medal at the World Championships in the individual all-around while also topping the USA in all-around score in qualifying cements his spot as our top all-around athlete. The second spot would go to either Asher Hong or Khoi Young. Because Asher Hong won the National Championship and was chosen above Young in the World Championships, the edge goes to him. 

Hong and Richard are both skillful at four of the six events, but neither covers the base of horizontal bar, so selecting an athlete for that event is important. The best option for this event is Paul Juda who made it to the horizontal bar finals round at the World Championships. Juda has also shown high levels of success in the floor exercise (won the National Championship) and the vault (made the World Championship final) making him an ideal third choice for the Olympic team. Khoi Young is the next best option to make the team. Not only did he place second in the individual all-around at the National Championshipms exemplifying his strength at all events, but he also got a silver medal in both the vault and pommel horse at the World Championships. Given the results of the National Championships, two men are in consideration for the final spot: Donnell Whittenberg and Yul Moldauer. Both placed first in one event at the National Championships and third in another The slight edge, however, goes to Whittenberg, who placed first in the rings at the National Championships, and this is the only event in which the USA did not send a single athlete to the finals of the World championships. While Moldauer has Olympic experience, Wittenberg’s expertise in the rings gives him a slight edge in our eyes. 

That gives a final five-man team of Frederick Richard, Asher Hong, Paul Juda, Khoi Young, and Donnell Whittenberg.

### Limitations

There were some limitations identified in our statistical methods. First, we arbitrarily assigned weights to each score and it is possible that different weightings could result in the selection of a different team. However, the weights used were reasonable given how recent performances are more likely to affect future performance and the desire to pick ‘clutch’ gymnasts. Another potential limitation is how we determined projected scores for athletes that have not competed in some time. Take Sunisa Lee as an example. Due to competing in college and injuries, Lee has very few scores since her Olympic performance in 2021. Because those performances account for many of her competitions, her lower recent scores are overshadowed by her higher older scores. Even with our weighting scheme her older scores account for a much greater proportion of her data. In this vein, athletes may have projected scores that do not resemble what their recent form might be. Some alternatives for our methodology could include using a Kernel Density Estimation method to estimate the distribution and then sampling, or obtaining all possible combinations of five athletes and then running an Olympics simulation. We could also introduce more factors to weighting such as the age of the competitor, skill level of the competition, etc.

## Appendix

Below are the graphs which validate the rough normality condition needed to generate t-distributions.

```{r, echo=FALSE, results = FALSE, warning= FALSE, message=FALSE}

data_2022_2023_f <- data_2022_2023 |>
  filter(Gender == "w")

data_2022_2023_m <- data_2022_2023 |>
  filter(Gender == "m")

usa_data_2022_2023_f <-  data_2022_2023_f |>
  filter(Country== "USA")

usa_data_2022_2023_m <- data_2022_2023_m |>
  filter(Country== "USA")

ggplot(aes(x= Score), data = usa_data_2022_2023_f) + geom_histogram() + facet_wrap(~Apparatus, scales = "free_y", ncol = 2) + labs(title = "Distribution of Scores for USA Women for Different Apparatuses", y = "Count")

ggplot(aes(x= Score), data = usa_data_2022_2023_m) + geom_histogram() + facet_wrap(~Apparatus, scales = "free_y", ncol = 2) + labs(title = "Distribution of Scores for USA Men for Different Apparatuses", y = "Count")


ggplot(aes(x= Score), data = data_2022_2023_f) + geom_histogram() + facet_wrap(~Apparatus, scales = "free_y", ncol = 2) + labs(title = "Distribution of Scores for All Women for Different Apparatuses", y = "Count")

ggplot(aes(x= Score), data = data_2022_2023_m) + geom_histogram() + facet_wrap(~Apparatus, scales = "free_y", ncol = 2)+ labs(title = "Distribution of Scores for All Men for Different Apparatuses", y = "Count")
```

\newpage

The following tables display the expected medal count for each US gymnast in each event, including in the all-around competition.

```{r, echo=FALSE, warning= FALSE, message=FALSE}
f_probs <- read_csv('App-1/data/expected_medals/f_total_probabilities.csv')

colnames(f_probs) <- c("Name", "AA", "BB", "FX", "VT", "UB")

f_probs |>
  kable()
```
\newpage
```{r, echo=FALSE, warning= FALSE, message=FALSE}
m_probs <- read_csv('App-1/data/expected_medals/m_total_probabilities.csv')

colnames(m_probs) <- c("Name", "AA", "SR", "FX","PB", "PH", "VT", "HB")

m_probs |>
  kable()
```

\newpage
### Works Cited
Brody Malone eyes spring 2024 return to gymnastics from leg surgeries. 17 May 2023. NBC Sports. https://www.nbcsports.com/olympics/news/brody-malone-gymnastics-injury-comeback. Accessed 20 Nov. 2023.


Duffy, P. Konnor McClain enrolling at LSU this fall with hopes of balancing NCAA and elite - Gymnastics Now. Gymnastics-Now.com. 13 Jul. 2023. https://gymnastics-now.com/konnor-mcclain-enrolling-at-lsu-this-fall-with-hopes-of-balancing-ncaa-and-elite/. Accessed 20 Nov. 2023.

Gunston, Jo. “Judging the Judges – How Statistical Analysis Evaluates Fairness And ...” Olympics.Com, 19 Oct. 2023, olympics.com/en/news/how-statistical-analysis-evaluates-fairness-accuracy-gymnastics. Accessed 06 Nov. 2023.


Most popular sports in the Summer Olympics US 2021. (n.d.). Statista. https://www.statista.com/statistics/1245746/summer-olympics-most-followed-sports-us/. Accessed 20 Nov. 2023.


Olympic Summer Games: global broadcast audience. (n.d.). Statista. https://www.statista.com/statistics/280502/total-number-of-tv-viewers-of-olympic-summer-games-worldwide/#:~:text=Between%202012%20and%202020%2C%20the,3.2%20billion%20five%20years%20earlier. Accessed 20 Nov. 2023.


Olympic champ Sunisa Lee gained 45 pounds due to kidney issue. “It was so scary.” 17 Nov. 2023. USA TODAY. https://www.usatoday.com/story/sports/olympics/2023/11/17/sunisa-lee-olympic-champion-kidney-health-paris/71616483007/. Accessed 20 Nov. 2023.


Tracy, Jeff. “Poll: Gymnastics Is Americans’ Most Highly Anticipated Olympic Sport.” Axios, Morning Consult, 19 July 2021, www.axios.com/2021/07/19/olympics-favorite-sport-gymnastics. Accessed 04 Dec. 2023. 

